---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
### Loading the data into R
```{r load}
read.csv("activity.csv") -> dat
```

## What is mean total number of steps taken per day?
### The mean and median steps/day are calculated below
```{r total steps}
aggregate(steps ~ date, dat, sum, rm.na = TRUE) -> per_day
hist(per_day$steps, main = "Steps Per Day", xlab = "steps")
mean(per_day$steps)
median(per_day$steps)
```


## What is the average daily activity pattern?
### The average pattern starts and ends with no activity, with a peak around interval 835.
```{r daily activity pattern}
aggregate(steps ~ interval, dat, mean, rm.na = TRUE) -> per_interval
plot(per_interval$interval, per_interval$steps, type = "l", xlab = "interval", ylab = "average steps")
per_interval[which.max(per_interval$steps),]
```


## Imputing missing values
### Missing values are imputed by replacing them with the average of that interval across all days. This increases the total daily steps but doesn't largely impact the mean/median daily steps
``` {r missing values}
library(dplyr)
inner_join(dat, per_interval, by = "interval") -> join
mutate(join, fullsteps = ifelse(is.na(steps.x), steps.y, steps.x)) -> dat_full
aggregate(fullsteps ~ date, dat_full, sum) -> per_day_full
hist(per_day_full$fullsteps)
mean(per_day_full$fullsteps)
median(per_day_full$fullsteps)
```
## Are there differences in activity patterns between weekdays and weekends?
### There is more activity throughout the day on weekends
``` {r weekdays and weekends}
mutate(dat_full, week = ifelse(weekdays(as.POSIXct(dat$date)) == "Saturday" | weekdays(as.POSIXct(dat$date)) == "Sunday", "Weekend", "Weekday")) -> dat_days
subset(dat_days, week == "Weekday") -> dat_wkday
aggregate(fullsteps ~ interval, dat_wkday, mean) -> per_day_full_weekday
subset(dat_days, week == "Weekend") -> dat_wkend
aggregate(fullsteps ~ interval, dat_wkend, mean) -> per_day_full_weekend
par(mfrow = c(2,1), mar = c(4,4,2,1))
with(per_day_full_weekday, plot(interval, fullsteps, type = "l", main = "Weekday"))
with(per_day_full_weekend, plot(interval, fullsteps, type = "l", main = "Weekend"))
```
